### 2022-03-01 paper/paper.tex
intro

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\fancyhf{}
\rhead{Uncertainty in college admissions markets \\~}
\lhead{Max Kapur \\ \date{\today}}
\cfoot{\thepage}

@@

\ifEN \section{Introduction}  \else \section{서론} \fi
This paper considers the optimization problem
%\begin{definition}[Optimal college application portfolio] \label{generalproblemstatement}
\begin{align}
\begin{split}
\text{maximize}\quad & \operatorname{E}\left[\max\bigr\{t_0,
\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\right] \\
\text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~\sum_{j\in \mathcal{X}} g_j \leq H
\end{split}
\end{align}
%\end{definition}
where $\mathcal{C} = \{ 1 \dots m\}$ is an index set, $H$ is a budget parameter, and for $j = 1 \dots m$, $g_j > 0$ is a cost parameter, $Z_j$ is a random, independent Bernoulli variable with probability $f_j$, and $t_j  > 0$ is a utility parameter. Without loss of generality, we assume that the $t_j$-values are weakly increasing in $j$, that $0 \leq t_0 \leq t_1$, and that each $g_j \leq H$.

We refer to this problem as the \emph{optimal college application portfolio,} as follows: Consider a college admissions market with $m$ college. The $j$th college is named $c_j$. Consider a single student, and let $f_j$ denote her probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not. It is appropriate to assume that the $Z_j$ are probabilistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates). Each $t_j$-value indicates the utility associated with attending $c_j$, and $t_0$ gives the utility associated with not attending college, an option named $c_0$. The application fee for $c_j$ is $g_j$, and $H$ represents the student's total budget to spend on application fees. Then the student's objective is to maximize the expected utility associated with the best school she gets into, and her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which the student applies.

As Chao (2014) remarked, this represents a somewhat subtle portfolio optimization problem. Traditional portfolio optimization models weigh the expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Meucci 2005). But college applicants maximize the observed value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent as to whether she gets into her $(j+1)$th choice. As a result, the valuation function that students maximize is \emph{convex} in the expected utility associated with individual applications. Risk management is implicit in the college application problem because, in a typical admissions market, college preferability is negatively correlated with competitiveness. That is, students  negotiate a tradeoff between highly attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet. Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization. Chao estimated her model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is possible. Our study pursues a more general solution.

Our study takes special interest in the validity of greedy solution algorithms, such as those that iteratively add the asset that yields the greatest increase in the objective function until the budget is exhausted. For certain classes of optimization problems, such as maximizing a submodular set function over a cardinality constraint, the greedy algorithm is known to be a good approximate solution and exact under certain additional assumptions (Fisher et al. 1978). For other problems, most notably the binary knapsack problem, the most obvious greedy algorithm can be made to perform arbitrarily poorly (Vazirani 2001).

We show analogous results for the college application problem: In the special case where each $g_j = 1$, the optimal portfolio satisfy a nestedness property that is equivalent to the validity of the greedy algorithm. As Rozanov and Tamir (2020) remark, the nestedness property is useful not only for obtaining solution algorithms, but in the implementation of an optimal policy under uncertain information. In the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges increasingly expect students to tailor their essays to the target school. Therefore, students may not know how many schools they can afford to apply to (in terms of both time and monetary costs) until late October. The nestedness property implies that students can prioritize their application strategy by working on applications in the order that they appear in the sequence of optimal application portfolios.

Unfortunately, the nestedness property does not hold in the general case, nor does the greedy algorithm offers any performance guarantee. We offer an alternative approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in fully polynomial time.

@@

\section{Notation and preliminaries} \label{preliminaries}
Before discussing the solution algorithms, we will introduce some additional notation and a few preliminary results that will come in handy.

@@

\section{(WIP) Numerical experiments} \label{numericalexperiments}
In this section, we present the results of numerical experiments designed to confirm the time complexity results established above. In both experiments, markets were generated by drawing $t_j$ independently from an exponential distribution with a scale parameter of ten and rounding up to the nearest integer. To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10 + Q)$, where $Q$ is a uniform draw from interval $[0, 1)$. In Experiment 1, which concerns Algorithm \ref{algorithmforlargeh}, we take each $g_j = 1$ and set $H = h = \lfloor m/ 2 \rfloor$. In Experiment 2, which concerns Algorithms \ref{ellisDP1} and \ref{ellisDP3}, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we take $H = \lfloor \frac{1}{2} \sum g_j \rfloor$. At each combination of the experimental variables, we generated 50 markets, and each computation was repeated three times, with the fastest of the three  recorded as the computation time. Therefore, each cell of each table represents 150 computations. We report the mean and standard deviation across the 50 markets. Where applicable, we do not count the time required to sort the entries of $t$.

>>>>

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\fancyhf{}
\rhead{How to apply to college \\~}
\lhead{Max Kapur \\ \date{\today}}
\cfoot{\thepage}

@@

\ifEN \section{Introduction}  \else \section{서론} \fi
This paper considers the portfolio optimization problem
%\begin{definition}[Optimal college application portfolio] \label{generalproblemstatement}
\begin{align}
\begin{split}
\text{maximize}\quad & \operatorname{E}\Bigl[\max\bigr\{t_0,
\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
\text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~\sum_{j\in \mathcal{X}} g_j \leq H
\end{split}
\end{align}
%\end{definition}
where $\mathcal{C} = \{ 1 \dots m\}$ is an index set, $H$ is a budget parameter, and for $j = 1 \dots m$, $g_j > 0$ is a cost parameter, $Z_j$ is a random, independent Bernoulli variable with probability $f_j \in (0, 1]$, and $t_j  > 0$ is a utility parameter. Without loss of generality, we assume that the $t_j$-values are weakly increasing in $j$, that $0 \leq t_0 \leq t_1$, and that each $g_j \leq H$.

We refer to this problem as the \emph{optimal college application portfolio,} as follows: Consider a college market with $m$ colleges. The $j$th college is named $c_j$. Consider a single prospective student in this market, and let each $t_j$-value indicate the utility she associates with attending $c_j$, where her utility is $t_0$ if she does not attend college. Let $g_j$ denote the application fee for $c_j$ and $H$ the student's total budget to spend on application fees. Finally, let $f_j$ denote her probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not. It is appropriate to assume that the $Z_j$ are probabilistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates). Then the student's objective is to maximize the expected utility associated with the best school she gets into within this budget. Her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies.

As Chao (2014) remarked, college application represents a somewhat subtle portfolio optimization problem. Traditional portfolio optimization models weigh the expected profit across all assets against a risk term, yielding a concave maximization problem with linear constraints (Meucci 2005). But college applicants maximize the observed value of their \emph{best} asset: If a student is admitted to her $j$th choice, then she is indifferent as to whether she gets into her $(j+1)$th choice. As a result, the valuation function that students maximize is \emph{convex} in the expected utility associated with individual applications. Risk management is implicit in the college application problem because, in a typical admissions market, college preferability is negatively correlated with competitiveness. That is, students  negotiate a tradeoff between highly attractive, selective “reach schools” and less preferable “safety schools” where admission is a safer bet. Finally, the combinatorial nature of the college application problem makes it difficult to solve using the gradient-based techniques used in continuous portfolio optimization. Chao estimated her model (which considers application as a \emph{cost} rather than a constraint) by clustering the schools so that $m=8$, a scale at which enumeration is possible. Our study pursues a more general solution.

We take special interest in the validity of greedy solution algorithms, such as that that iteratively adds the asset that yields the greatest increase in the objective function until the budget is exhausted.
% Greedy algorithms produce a nested, parametric family of solutions. As Rozanov ...
For certain classes of optimization problems, such as maximizing a submodular set function over a cardinality constraint, the greedy algorithm is known to be a good approximate solution and exact under certain additional assumptions (Fisher et al. 1978). For other problems, notably the binary knapsack problem, the most intuitive greedy algorithm can be made to perform arbitrarily poorly (Vazirani 2001).

We show analogous results for the college application problem: In the special case where each $g_j = 1$, the optimal portfolio satisfy a nestedness property that is equivalent to the validity of the greedy algorithm. As Rozanov and Tamir (2020) remark, the nestedness property is useful not only for obtaining an optimal solution, but in the implementation of an optimal policy under uncertain information. In the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges increasingly expect students to tailor their essays to the target school. Therefore, students may not know how many schools they can afford to apply to (in terms of both time and monetary costs) until late October. The nestedness property implies that students can prioritize their application strategy by working on applications in the order that they appear in the sequence of optimal application portfolios.

Unfortunately, the nestedness property does not hold in the general case, nor does the greedy algorithm offers any performance guarantee. Instead, we offer a pseudopolynomial-time algorithm that is tractable for typical college market instances, as well as an approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in fully polynomial time.

@@

\section{Notation and preliminary results} \label{preliminaries}
Before discussing the solution algorithms, we will introduce some additional notation and a few preliminary results that will come in handy.

@@

\section{(WIP) Numerical experiments} \label{numericalexperiments}
In this section, we present the results of numerical experiments designed to confirm the time complexity results established above. In both experiments, markets were generated by drawing $t_j$ independently from an exponential distribution with a scale parameter of ten and rounding up to the nearest integer. To achieve partial negative correlation between $t_j$ and $f_j$, we then set $f_j = 1 / (t_j + 10 + Q)$, where $Q$ is drawn uniformly from the interval $[0, 1)$. In Experiment 1, which concerns Algorithm \ref{algorithmforlargeh}, we take each $g_j = 1$ and set $H = h = \lfloor m/ 2 \rfloor$. In Experiment 2, which concerns Algorithms \ref{ellisDP1} and \ref{ellisDP3}, each $g_j$ is drawn uniformly from the set $\{5, \dots, 10\}$ and we take $H = \lfloor \frac{1}{2} \sum g_j \rfloor$. At each combination of the experimental variables, we generated 50 markets, and each computation was repeated three times, with the fastest of the three  recorded as the computation time. Therefore, each cell of each table represents 150 computations. We report the mean and standard deviation across the 50 markets. Where applicable, we do not count the time required to sort the entries of $t$.
